Apache Flink\texttrademark : Streaming Dataflows as a Basis \\ for Universal Data Analytics

### 2015-11-10 architecture.tex
Figures and listings update

\begin{figure}[h!]
	\centering  	
  	\includegraphics[width=.55\textwidth]{figs/stack.pdf}
	\vspace{-2mm}
	\caption{The Flink software stack.}
	\vspace{-3mm}
	\label{fig:FlinkStack}
\end{figure}

\para{Flink's Runtime and APIS.} The core of Flink is the distributed dataflow engine, which executes dataflow programs called Job Graphs. A Job Graph is a DAG of operators and connections between operators. There are two ``core'' APIs in Flink: the DataSet API for processing finite data sets (often referred to as ``batch processing''), and the DataStream API for processing potentially unbounded data streams (often referred to as ``stream processing''). Flink's core runtime engine can be seen as a streaming dataflow engine, and both the DataSet and DataStream APIs create programs (Job Graphs) accepted by this engine. On top of the core APIs, Flink bundles ``domain-specific'' libraries and APIs that generate DataSet and DataStream API programs. Currently these are the following:
\begin{inparaenum}[i)]
  \item Gelly, an API and library for processing graphs,
  \item FlinkML, an API and library for composing Machine Learning pipelines and
  \item Table, an API similar in spirit to Microsoft's LINQ.
\end{inparaenum}
A Flink cluster comprises of three types of processes: the client, the JobManager, and the TaskManager. The client takes the program code (written in a mixture of Flink's APIs), transforms it to a Job Graph, and submits it to the JobManager. The compilation process involves a type extraction and checking phase that generates serializers and comparators for all used types. DataSet programs, also go through a cost-based query optimization phase, similar to the physical optimizations performed by relational query optimizers (more details in Section X).

@@

>>>>

\para{Flink's Runtime and APIS.} The core of Flink is the distributed dataflow engine, which executes dataflow programs called Job Graphs. A Job Graph is a DAG of operators and connections between operators. There are two ``core'' APIs in Flink: the DataSet API for processing finite data sets (often referred to as ``batch processing''), and the DataStream API for processing potentially unbounded data streams (often referred to as ``stream processing''). Flink's core runtime engine can be seen as a streaming dataflow engine, and both the DataSet and DataStream APIs create programs (Job Graphs) accepted by this engine. On top of the core APIs, Flink bundles ``domain-specific'' libraries and APIs that generate DataSet and DataStream API programs. Currently these are the following:
\begin{inparaenum}[i)]
  \item Gelly, an API and library for processing graphs,
  \item FlinkML, an API and library for composing Machine Learning pipelines and
  \item Table, an API similar in spirit to Microsoft's LINQ.
\end{inparaenum}
A Flink cluster comprises of three types of processes: the client, the JobManager, and the TaskManager. The client takes the program code (written in a mixture of Flink's APIs), transforms it to a Job Graph, and submits it to the JobManager. The compilation process involves a type extraction and checking phase that generates serializers and comparators for all used types. DataSet programs, also go through a cost-based query optimization phase, similar to the physical optimizations performed by relational query optimizers (more details in~\autoref{sec:batch}).

@@

\begin{figure}[ht]
\centering     %%% not \center
\subfigure[The Flink software stack.]{\label{fig:FlinkStack}\includegraphics[width=.5\textwidth]{figs/stack.pdf}}
\subfigure[The JobGraph is the logical view of a Flink job.]{\label{fig:JobGraph}\begin{tikzpicture}
[shorten >=1pt,node distance=1.75cm,>=stealth',initial text=, font=\tiny]
\tikzstyle{op}=[circle, draw=blue!50,very thick,fill=blue!20, text width=0.5cm]
\tikzstyle{ir}=[cylinder, cylinder uses custom fill, cylinder body fill=orange!50, cylinder end fill=orange!50, shape border rotate=90, aspect=0.35, draw, minimum size=1cm, text width=0.5cm, align=center]

\node[op] (op_1) {~~~~~~~};
\node[ir] (ir_1) [right of = op_1] {~~~~~~~};
\node[op] (op_2) [right of = ir_1] {~~~~~~~};
\node[ir] (ir_2) [right of = op_2] {~~~~~~~};
\node[op] (op_3) [below of = op_2] {~~~~~~~};
\node[ir] (ir_3) [right of = op_3] {~~~~~~~};
\node[op] (op_4) [right of = ir_2] {~~~~~~~};
\node[op] (op_5) [right of = ir_3] {~~~~~~~};

\path[->] 	
			(op_1) 	edge node [above] {} (ir_1)
			(ir_1) 	edge node [above] {} (op_2)
			(op_2) 	edge node [above] {} (ir_2)
			(ir_2) 	edge node [above] {} (op_4)
			(op_3) 	edge node [above] {} (ir_3)
			(ir_3) 	edge node [above] {} (op_4)
			(ir_3) 	edge node [above] {} (op_5)
			;
\end{tikzpicture}}
\caption{Apache Flink architecture.}
\end{figure}

### 2015-11-10 exec-model.tex
Figures and listings update

\section{Execution model}
\label{sec:execution}
\para{The Job Graph.} Flink's execution model is based on the Job Graph, a directed acyclic graph (DAG) that consists of nodes and edges. There are two classes of nodes: (stateful) operators, and (logical) intermediate results (IRs). For example, the graph below consists of five operators (circles), and three intermediate results. Operators abstract computation (e.g., transformations, joins, etc), state (e.g., a persistent counter), as well as data sources (e.g. reading data from a file system, a socket, a message queue, etc.), and data sinks. Operators produce intermediate results, as well as updates to state. An intermediate result is a logical handle (pointer) to the data that is produced by one operator. An intermediate result can be consumed by one or more operators. Intermediate results are logical in the sense that the data they point to may or may not be materialized on disk. When the JobGraph is scheduled in a cluster for execution, it is parallelized to form an ExecutionGraph that consists of tasks (parallel instances of operators), and intermediate result partitions (IRPs).

\begin{figure}[ht]
\centering
\begin{tikzpicture}
[shorten >=1pt,node distance=1.75cm,>=stealth',initial text=, font=\tiny]
\tikzstyle{op}=[circle, draw=blue!50,very thick,fill=blue!20]
\tikzstyle{ir}=[cylinder, cylinder uses custom fill, cylinder body fill=orange!50, cylinder end fill=orange!50, shape border rotate=90, aspect=0.35, draw, minimum size=1cm, text width=0.5cm, align=center]

\node[op] (op_1) {~~~~~~~};
\node[ir] (ir_1) [right of = op_1] {~~~~~~~};
\node[op] (op_2) [right of = ir_1] {~~~~~~~};
\node[ir] (ir_2) [right of = op_2] {~~~~~~~};
\node[op] (op_3) [below of = op_2] {~~~~~~~};
\node[ir] (ir_3) [right of = op_3] {~~~~~~~};
\node[op] (op_4) [right of = ir_2] {~~~~~~~};
\node[op] (op_5) [right of = ir_3] {~~~~~~~};

\path[->] 	
			(op_1) 	edge node [above] {} (ir_1)
			(ir_1) 	edge node [above] {} (op_2)
			(op_2) 	edge node [above] {} (ir_2)
			(ir_2) 	edge node [above] {} (op_4)
			(op_3) 	edge node [above] {} (ir_3)
			(ir_3) 	edge node [above] {} (op_4)
			(ir_3) 	edge node [above] {} (op_5)
			;
\end{tikzpicture}
\label{fig:JobGraph}
\caption{The JobGraph is the logical view of a Flink job.}
\end{figure}

>>>>

\section{Execution model}
\label{sec:execution}
\para{The Job Graph.} Flink's execution model is based on the Job Graph, a directed acyclic graph (DAG) that consists of nodes and edges. There are two classes of nodes: (stateful) operators, and (logical) intermediate results (IRs) as presented on Figure \autoref{fig:JobGraph}. For example, the aforementioned graph consists of five operators (circles), and three intermediate results (cylinders). Operators abstract computation (e.g., transformations, joins, etc), state (e.g., a persistent counter), as well as data sources (e.g. reading data from a file system, a socket, a message queue, etc.), and data sinks. Operators produce intermediate results, as well as updates to state. An intermediate result is a logical handle (pointer) to the data that is produced by one operator. An intermediate result can be consumed by one or more operators. Intermediate results are logical in the sense that the data they point to may or may not be materialized on disk. When the JobGraph is scheduled in a cluster for execution, it is parallelized to form an ExecutionGraph that consists of tasks (parallel instances of operators), and intermediate result partitions (IRPs).
